---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

# title: "Helping You Write Academic Papers in R using Texevier"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: TRUE
Entry1: "A Game Theoretic Approach to Deadline Adherence"
Entry2: "Microeconomics 871 Essay" 
Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
Logo_width: 0.3 # If using a logo - use this to set width (size) of image
Entry3: "\\textbf{Jessica van der Berg - 20190565}"
Entry4: "\\textbf{Laura Meyer - 20748302}"
Entry5: "\\textbf{Cassandra Pengelly - 20346212}"
Entry6: "18 October 2021 | Word Count: 1980"
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
# AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
# Author1: "Nico Katzke^[__Contributions:__  \\newline _The authors would like to thank no institution for money donated to this project. Thank you sincerely._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
# Ref1: "Prescient Securities, Cape Town, South Africa" # First Author's Affiliation
# Email1: "nfkatzke\\@gmail.com" # First Author's Email address
# 
# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: TRUE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
# abstract: |
  # Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
```

<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->
\newpage

# Introduction \label{intro} 

The relationship between a supervisor and a subordinate has been studied extensively in the social sciences [@power], [ @comm], [@trust] and it's been found that a subordinate’s action is often influenced by the behaviour of the supervisor. Identity-based motivation theory posits that a relationship of trust and leniency between a subordinate and a supervisor will ultimately lead to both parties being better off in terms of reaching their goals, performance, and mental health [@trust]. Game theory models provide a useful framework to analyze the interactions between a supervisor and a subordinate [@book p.1]. A common situation that arises is a subordinate being assigned a task that should be complete before a deadline. Game theory can offer insights as to when a subordinate should submit a task on time or miss the deadline if there is incomplete information.

This article contributes to the literature on games of incomplete information, and dynamic interactions between students and lecturers. We investigate a case where a student is required to submit an assignment with by a certain deadline, but she experiences a crisis and must choose whether to submit her assignment on time or late. We impose a structure of continuous types for both players, with a discrete set of actions. This essay is organised as follows: section \ref{lit} briefly discusses the literature on games of incomplete information and some applications. Section \ref{game} presents our model of deadline adherence; and section \ref{result} analyses the results of the game. Section \ref{extension} provides an extension of the game and the final section concludes \ref{con}.

# Games of Incomplete Information \label{lit} 

In real-life situations, we often don't have full information before we make decisions @2020games. @von [p.30] first used the term \emph{incomplete information} in reference to a game theory model in which parts of the normal form structure are unspecified. However, Von Neumann and Morgenstern deemed further research into such a model as unimportant [@2004com]. @luce1956 disagreed, and extended on the incomplete information literature by assuming that each player has a perception^[This belief does not necessarily need to be correct] of the payoff function of the other player. However, games of incomplete information were practically difficult to solve until @harsanyi developed a general analytical framework. 

We follow the approach of @harsanyi in applying a game-theoretic model of incomplete information, where players have less than full information about each others payoff functions. Based on the Bayesian methodology, both players have expectations in the form of subjective probability distributions. Players have different types, which are randomly assigned and represent their belief about the game being played. However, players do not know the type of the other player [@2020games]. Both players attempt to estimate the probability of each others types, subject to the available information. To solve the outcome of the model, the game of incomplete information will be reinterpreted as a game with complete and imperfect information, by transforming its basic mathematical structure.

# A Model of Deadline Adherence \label{game}

This section presents a model of deadline adherence, with the purpose of understanding how a student thinks about assignment submissions, and how a lecturer acts when submissions are late. An explanation of the model follows. A student receives an assignment, which is due by a certain date set by the lecturer. While the student is working on the assignment, she experi a crisis and therefore spends less time on the assignment. She has two options: she can hand in the assignment on time or she can hand in late. If she hands in on time, she will get a payoff of $a-c$, where $a$ is her potential pre-crisis mark, and $c$ is the negative impact the crisis has on her mark. However, if she submits her assignment late, she has some time to recover after the crisis and reduce its academic impact. Her payoff is $a-\beta c -m$ if the lecturer gives her a penalty, where $m$ is the size of the penalty. She gets a payoff of $a-\beta c$ if there is no penalty. $\beta$ represents the type of the student, where a high $\beta$ suggests a low resiliency to crises, and a low $\beta$ suggests a high resiliency and a better academic recovery. The student observes her own type but does not know the lecturer's type.

On the other hand, the lecturer is faced with the decision either to give a penalty ($m$), if a student submits late, or not to give a penalty. If the lecturer gives a penalty, he feels bad since the student has gone through a crisis. The size of his disutility depends on the size of the penalty ($m$) and how empathetic the lecturer is, where the level of empathy describes the lecturer's type ($\delta$). The more empathetic the lecturer is, the higher $\delta$ is. The lecturer observes his own type but not that of the student. The lecturer's and student's types are both continuous types, which are independently and randomly chosen by nature at the start of the game from a uniform distribution^[A uniform distribution puts equal chance on any of the outcomes between 0 and 1 happening.]: $\delta \sim Uniform(0,1)$ and $\beta \sim Uniform(0,1)$. If the lecturer decides not to impose a penalty, he feels good that he did not impose on a student experiencing a crisis, and gets a positive payoff of $\delta c$. However, the lecturer knows that by waving the penalty, he may be encouraging this student, and other students to hand in late in the future. The lecturer would rather deter late hand-ins, and receives a negative payoff $-d$ for not deterring^[This deterrent parameter relates to the literature on games of repeated interaction and reputations [@deter].] late hand-ins.

The parameters $a,\, c,\, m\, \&\, d$ are all common knowledge. This is a game of incomplete information because the players' types are not common knowledge. The type spaces are continuous and the action spaces are discrete. Each player needs to choose his/her action based on his/her own type, what each believes the other player's type is, and the values of $a,\, c,\, m\, \&\, d$. Figure \ref{Figure1} shows the game in extensive form^[The simultaneous form game can be found in the appendix, \ref{tab1}]. And a summary of the game's parameters and restrictions are given in figure \ref{sum} below.

\includegraphics[scale=1]{"img/continuous.jpg"}
\captionof{figure}{This game is dynamic, where nature first chooses the student's and lecturer's types. The dashed line represents information only known by the lecturer and the solid line represents information only known by the student. Then the student moves, deciding to hand in on time or late after experiencing a crisis. If the student hands in late, the lecturer decides to impose a penalty or not. }
\label{Figure1}


```{r sum, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}

library(xtable)
source("code/game_summary.R")
data<- game_summary()


table <- xtable(data, caption = "Game Parameters \\label{sum}")
  print.xtable(table,
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H',
             # scalebox = 0.3,
             comment = FALSE,
             caption.placement = 'bottom',
             include.rownames = FALSE,
             sanitize.text.function=function(x){x},
             booktabs = T # Aesthetics
             )

```

# Results and Discussion \label{result}
 
In order to understand how the lecturer and student will make their decisions given their beliefs, we need to solve for their best responses. A best response for the student would be to hand in on time if the expected payoff from handing in on time is higher than the expected payoff of submitting late. Defining $p$ as the probability that the lecturer will give a penalty, a student should hand in on time where: 
\begin{align*}
\beta>\frac{c-m p}{c}
\end{align*} 
The right hand side is a constant^[Since $c$ and $m$ are known, and $p$ is a belief the student holds.]. This implies that there is some threshold value of $\beta$, for which a student should hand in on time. If the student believes that the lecturer will give no penalty (i.e. $p=0$), then she should only hand in on time if $\beta > \frac{c}{c} \Rightarrow \beta > 1$. Since $\beta$ lies between $0$ and $1$, the inequality will never hold and she should always hand in late. From an intuitive stand point, this makes sense: a student can never do worse by handing in late if there is no penalty^[Handing in late is a weakly dominant strategy.] but she will do better to hand in late if she is resilient in any way ($\beta < 1$) and can partially recover from the crisis.

However, if the student believes that the lecturer will give a penalty with some positive probability ($p>0$) then her decision to hand in on time depends on her level of resiliency, the magnitude of the crisis and the  size of the penalty. As the cost of the crisis increases, the threshold value increases (ceteris paribus), and the student becomes more likely to hand in late (unless her she has a very low resiliency). If the mark penalty is high, the threshold value is smaller, and the student is more likely to hand in on time (unless is very resilient: $\beta$ is very low). We can analyse the lecturer's best response rule similarly. The lecturer should impose a penalty where:  
\begin{align*}
\delta<\frac{d}{c+m}
\end{align*}
The threshold value for the lecturer is the ratio between the deterrent factor and the sum of the crisis cost and the penalty mark. If the deterrent factor is high relative to the cost of crisis and the penalty mark, then the lecturer is more likely to give a penalty (unless he is highly empathetic). If the cost of the crisis large and the mark penalty is high relative to the deterrent factor, the lecturer is more likely to waive the penalty (unless he has very low empathy levels). 

While the best response analysis is useful, if we want to understand the outcome of the game and the players' strategies, we need to solve for the Bayesian Nash Equilibrium (BNE). The full derivations of the solution concepts are given in the appendix (\ref{B}).

After observing her private type, the student chooses the following hand-in pattern:  
\begin{align*}
s_{\beta}^{*}(\beta)=\left\{\begin{array}{lll}
\text{On time} & \text{if} & \beta>1-\frac{m d}{c^{2}+m c}  \\
\text{Late} & \text{if} & \beta \leq 1-\frac{m d}{c^{2}+m c} 
\end{array}\right.
\end{align*}
 
And the lecturer's BNE strategy is given by: 
\begin{align*}
s_{\delta}^{*}(\delta)=\left\{\begin{array}{lll}
\text{Penalty} & \text { if } & \delta<\frac{d}{c+m} \\
\text{No penalty} & \text { if } & \delta \geq \frac{d}{c+m}
\end{array}\right.
\end{align*}
  
When the lecturer and the student play their equilibrium strategies, neither has an incentive to deviate and we get a Bayesian Nash equilibrium. We would interpret the student's BNE strategy similarly to her best response; however, her strategy profile only depends on her own type and no longer on her beliefs about the lecturer's type. The lecturer's best response function and equilibrium strategy profile are the same because the lecturer observes whether the student hands in late or not and therefore has no need to hold beliefs about when the student will hand in.

# Extension \label{extension}

We extend the game explained in section \ref{game} to account for grade inflation. The lecturer has an incentive to inflate the marks of the student since a higher grade is associated with the student giving a positive evaluation of the lecturer which could lead to an increase in salary and promotions. However, inflation the mark if ethically wrong therefore the lecturer incurs cost \(\gamma\) if he decides to inflate. If he decides to leave the mark unchanged, then he occurs a sympathy cost \(\delta\). A student experiences a cost of \(\omega\) for asking the lecturer and the lecturer experiences a cost of \(\phi\) of being bothered by the lecturer. 
If the lecturer decides not to penalize the student, then he will never choose to inflate a students mark. Knowing this, a student that hands in late and receives no penalty will always accept his mark. If a student hands-on time, a lecture will always leave the mark unchanged at the request of the student, therefore the student will always accept the mark. 
If the lecturer decides to penalize a late student, the lecturer will choose to inflate a student’s mark if the moral cost is smaller than the empathy cost the lecturer experiences. 
\[\gamma < \delta m \]
If the moral cost that the lecturer experiences is bigger than the empathy cost the lecturer experiences for giving a penalty, then the lecturer is strict, otherwise the lecturer is lenient. Therefore, a strict lecturer will not inflate a student mark but a lenient lecture will. 

A summary of the game's parameter's and restrictions are shown in table \ref{table:ext} and the game is represented in Figure \ref{fig:extd}

\begin{table}
\caption{Extended Game Parameters}
\centering
\begin{tabular}{c c c} \hline
Parameter & Explanation & Restrictions \\
\hline 
\(\omega\) & Exogenous cost of asking for the lecturer for higher mark & 0 < \(\omega\) $\le$ 1 \\
\(\delta\) & Ethical cost of inflating students mark & 0 < \(\delta\) $\le$ 1 \\
\(\phi\) & Cost of annoyance lecture experience for student bothering him & 0 <\(\phi\) $\le$ 1 \\
x & Additional marks that student get when lecturer decides to inflates &  1 < x $\le$ 2 \\ [1ex]
\hline 
\end{tabular}
\label{table:ext}
\end{table}


\includegraphics[scale=0.90]{"img/extend.jpg"}
\captionof{figure}{This is an extended version of the game presented in Figure 3.1. To account for grade inflation, the student gets to accept the grade the lecturer game him, or reject the grade and ask the lecturer for a better mark. The lectuerer can then decided to inflate (I) the students mark or leave the mark unchanged (U). }
\label{fig:extd}


# Conclusion \label{con}
The results of the model are intuitive and provide a useful insight into how student's think about handing in assignments, and how lecturers respond to late submissions. One shortcoming of the model is its assumption that $c$ is common knowledge. Although some lecturers are in touch with their students and would know whether they are experiencing a crisis and how impactful the crisis would be, most lecturers have too many students to know that information.

Extensions, generality
Shortcomings

\newpage

# References {-}

<div id="refs"></div> 

\newpage
# Appendix A \label{A} {-}

```{r tab1, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}

library(xtable)
source("code/simul.R")
data<- simul()


table <- xtable(data, caption = "Strategic form of the game \\label{tab1}")
  print.xtable(table,
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H',
             # scalebox = 0.3,
             comment = FALSE,
             caption.placement = 'bottom',
             include.rownames = TRUE,
             sanitize.text.function=function(x){x},
             booktabs = T # Aesthetics
             )

```

# Appendix B \label{B} {-}

## Payoffs \label{payoff} {-}

Student payoffs:
\begin{align*}
E[\text{On Time}]&= a- c \\
E[\text{Late}]&=  p(a-\beta c-m) +(1-p)(a-\beta c) \\
&=-m p+a-\beta c
\end{align*}
Student plays on time if:
\begin{align*}
a-c>a-m p-\beta c \\
\beta c>c-m p \\
\beta>\frac{c-m p}{c}
\end{align*}
Student plays late if:
\begin{align*}
\beta<\frac{c-m p}{c}
\end{align*}
Lecturer Payoffs:
\begin{align*}
E[\text{Penalty}]&=q(-\delta m)+(1-q)(0) \\
&=q(-\delta m) \\
E[\text{No Penalty}] &=q(\delta c-d)+(1-a)(0) \\
&=q(\delta c-d)
\end{align*}
Lecturer gives a penalty if:
\begin{align*}
q(-\delta m)&>q(\delta c-d) \\
-\delta m&>\delta c-d \\
d&>\delta(c+m) \\
\delta&<\frac{d}{c+m} \\
\delta &<\bar{\delta}
\end{align*}
Lecturer gives no penalty if:
\begin{align*}
\delta &\geq \frac{d}{c+m} \\
\delta &\geq \bar{\delta} \\
\end{align*}

## Best Responses \label{br} {-}
Solving for the best responses:
\begin{align*}
p=\text{Probability that the lecturer gives a penalty} = \bar{\delta}=\operatorname{Prob}(\delta<\bar{\delta})
\end{align*}
Substitute into the student's best response function - student hands in on time if:
\begin{align*}{}
\beta>\frac{c-m(\bar{\delta})}{c}
\end{align*}{}
Since $0 \leq \beta \leq 1$, $\beta$ cannot be greater than 1. This implies
\begin{align*}{}
\frac{c-m(\bar{\delta})}{c} \leq 1 \\
c-m \bar{\delta} \leq c \\
-m \bar{\delta} \leq 0 \\
0 \leq \bar{\delta}
\end{align*}{}
Since $0 \leq \bar{\delta} \leq 1$, this condition will always hold.

$\beta$ cannot be less than 0:
\begin{align*}
\frac{c-m \bar{\delta}}{c}&<0 \\
c-m \bar{\delta}&<0 \\
-m \bar{\delta}&< -c \\
\bar{\delta}&>\frac{c}{m}
\end{align*}
if $\bar{\delta}>\frac{c}{m} \Rightarrow \beta=0$, otherwise:
\begin{align*}
\beta =\frac{c-m \bar{\delta}}{c}
\end{align*}
Best response function for the student:
\begin{align*}
B R_{\beta}(\bar{\delta})=\left\{\begin{array}{lll}
\frac{c-m\bar{\delta}}{c} & \text { if } & \bar{\delta}\leq \frac{c}{m} \\
0 & \text { if } & \bar{\delta}> \frac{c}{m}
\end{array}\right.
\end{align*}
Best response function for the lecturer:
\begin{align*}
B R_{\delta}(\delta)=\left\{\begin{array}{lll}
\text{Penalty} & \text { if } & \delta<\frac{d}{c+m} \\
\text{No penalty} & \text { if } & \delta \geq \frac{d}{c+m}
\end{array}\right.
\end{align*}

## Bayesian Nash Equilibrium \label{bay} {-}
The Bayesian Nash equilibrium occurs at the point where the best response functions intersect.
For the BRFs to cross:
\begin{align*}
\text{Substitute} \; \bar{\delta} &= \frac{d}{c+m} \; \text{into} \; \beta=\frac{c-m\bar{\delta}}{c} \\
\text{Then:} \; \beta&=\frac{c}{c}-\frac{m}{c}\left(\frac{d}{c+m}\right) \\
\beta&=1-\frac{m d}{c^{2}+m c} \\
\end{align*}
BNE strategy for the student
\begin{align*}
s_{\beta}^{*}(\beta)=\left\{\begin{array}{lll}
\text{On time} & \text{if} & \beta>1-\frac{m d}{c^{2}+m c} \\
\text{Late} & \text{if} & \beta \leq 1-\frac{m d}{c^{2}+m c}
\end{array}\right.
\end{align*}
BNE strategy for the lecturer
\begin{align*}
s_{\delta}^{*}(\delta)=\left\{\begin{array}{lll}
\text{Penalty} & \text { if } & \delta<\frac{d}{c+m} \\
\text{No penalty} & \text { if } & \delta \geq \frac{d}{c+m}
\end{array}\right.
\end{align*}



